Knowledge Representation for Service Robots: A Survey

### 2014-04-21 krs-survey.tex
reworked text, added references

\newcommand{\taxon}[2]{%
    \textbf{[#1] \emph{#2}}
}

@@

The aim of this contribution is two-fold: we first propose a typology of
knowledge representation systems for service robotics that covers intrinsic
features like expressiveness or reasoning techniques as well as extrinsic
features like knowledge acquisition and integration into pre-existent robotic
architectures. We base the typology on an initial typical scenario of service
robotics.

We then survey the current knowledge representation system in the community to
identify which knowledge-bound cognitive abilities are today well understood
and implemented, and which would require more research efforts.

We hope this article may bring a clearer picture of the challenges of knowledge
representation and manipulation for the robotic research community.

@@

\subsection{Knowledge as the Root of Cognitive Abilities}
\label{sect|cognitive-abilities}

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Foundations of Artificial
Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
wrote together in 1998:

@@

@@

    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.''

To be answered, those questions have in common one prerequisite: the robot must
be able to manipulate \emph{explicitly} knowledge, and hence, need a way to
represent the knowledge.

This survey focuses on this knowledge representation issue: we aims at first
establishing a comprehensive typology of representational needs for
robotics in the specific context of service robotics, and then at painting the
current landscape of approaches to the knowledge representation problem in the
research community, in order to summarize the major trends and to identify the
possible shortcomings.

We have classified the features of knowledge representation systems (KRS) into
six broad categories: \emph{what can be represented?} or the
\emph{expressiveness} question ; \emph{how things are represented?} ; \emph{what
reasoning techniques are offered?} ; \emph{how knowledge is acquired and
anchored?} ; \emph{how the knowledge is made available and used by the other
components of the robot?} and \emph{which actual knowledge is stored?} or the
\emph{knowledge instantiation} question.

We have also tried, for each project, to make clear \begin{inparaenum} \item
how informations from perception or interaction are turned into knowledge,
\item how, in return, symbolic concept are \emph{anchored} into the robot
sensori-motor space, \item how the knowledge representation system integrates
with the decisional layers of the robot, and how this can improve or ease
robotic control.\end{inparaenum}
\fxfatal{Do we really do it?}

\paragraph{What do we call ``knowledge''?}
\label{sect|on-knowledge}

Since we will discuss at length the concept of knowledge in the context of
robotics in the coming pages, it is useful to make our terminology explicit.

Be it in philosophy, cognitive sciences or computer sciences, reaching an
agreement on a definition of ``knowledge'' seems difficult.

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language, semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the article, give a second look to
this distinction.

In our robotic context, we define knowledge as a narrower concept, while
keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
interrelated logical facts that are meaningful to the robot executive
controller}. By \emph{meaningful} we mean that can possibly be interpreted to
lead to a purposeful action. We will see that our main challenge while
designing a cognitive architecture is furthermore to make this knowledge as
\emph{explicit} as possible.

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question. In this article,
we will associate data to low-level material like raw sensor output, and
information to uncontextualised symbolic facts.

To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, coords)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

\subsection*{Survey Inclusion Criteria}
\label{sect|inclusion-criteria}

Every robotic system has, implicitly or not, some knowledge representation
systems. It may range from a simple state vector to an explicit symbolic
knowledge base.  This survey focuses on the right end of this spectrum:
symbolic systems, suited for abstract reasoning.

Besides, we have decided to restraint the set of systems to those actually
implemented on robots, and used in semantic-rich environments (\ie dynamic,
partially unknown environments with a large range of different entities which
may have interactions). The typical scenario that would involve such robots is
a service robot in a human-friendly environment like a kitchen.

We have limited ourselves to systems that
\begin{inparaenum} 
    \item  run on \emph{service robot} (that is, robots that interact with 
    objects in a semantic-rich environment primarily designed for humans),
    \item  ground the knowledge in the physical world (physically embedded
    systems able to assess their environment),
    \item  are able to merge different knowledge modalities,
    \item  are able of on-line, dynamic knowledge acquisition and reasoning 
    (\ie not simple static databases).
\end{inparaenum}

These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.

We have also chosen not to include the GOLOG language and its
derivatives~\cite{levesque1997golog, Ferrein2008, Gspandl2011} in this survey. While
several implementations on robots, including service robots, do exist, the
focus of this language is on representation and reasoning about actions and
situations, and the link with symbolic, abstract knowledge is not explicit.

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
GLAIR~\cite{Shapiro2009} or {\sc ACT-R} have declarative knowledge
modules~\cite{Derbinsky2010} and have been recently used on service robots (see
{\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are also absent from this
survey because we did not find much references in the literature on knowledge
manipulation and representation applied to real-world robotic scenarii for
these architectures.

A comprehensive reference on (bio-inspired) cognitive architectures is also
available from the BICA Society~\cite{BicaCogArch2011}.

\subsection{Previous work}
\label{sect|evaluation-literature}

The typology we propose is in part based on a comprehensive synthesis of
classifications and analysis found in the literature. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation in the
context of service robotics.

Levesque and Lakemeyer~\cite{Levesque2008} present in their chapter on
Cognitive Robotics several characteristics of knowledge representation systems
for robots, stressing the need of representing the \emph{dynamics of the
world}.  Sensing is included in the knowledge representation via
\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
distinct parallel mental models; action representation and reasoning about
tasks is discussed in the context of \emph{situation calculus}; \emph{open
world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
robot programming and knowledge representation can be related. We integrate
most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
\emph{knowledge processing middleware} as systems supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context dependent processing at many different levels of abstraction''. They
identify six characteristics: the system must be able to \emph{merge
informations} from different, possibly distributed sources; it should support
quantitative as well as qualitative processing of information, it should offer
\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
(which require what we call here \emph{non-monotonicity}) and finally
\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
specification of the processing functionalities'').

Several surveys compare global cognitive architectures \cite{Langley2006,
Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
distinguish nine capabilities: recognition and categorisation, decision making,
perception and situation assessment, prediction and monitoring, planning,
reasoning, execution control, interaction and
learning/remembering/introspection. They also separately identify four
\emph{properties} of a cognitive architecture, that categorise how knowledge is
handled by the architecture: representation of knowledge, organisation of
knowledge, utilisation of knowledge and acquisition and refinement of
knowledge. This categorisation had a notable influence on our typology, and
many of these categories are also present in our proposal.

Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology}
(how the structure of the world is captured in a representation and shared),
the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.

At an even broader scope, several authors from fields that are connected to
robotics have previously listed desirable features of artificial systems aiming
at rich cognitive abilities.

For instance McCarthy recently listed in~\cite{McCarthy2007} the challenges he
identifies on the road to a \emph{human-level AI}.

	\item the ability to \emph{"operate successfully in the common sense
	informatic situation"},

	\item the necessity of relying on mathematical logic, as the most fruitful
	formalism for machine intelligence,

	\item the ability to deal with \emph{approximate concepts and approximate
	theories} (that would include representing them, and reasoning with them),

	\item non-monotonic reasoning,

	\item what McCarthy calls \emph{Elaboration Tolerance}: the ability to
	extend \emph{on demand} the closed domain of interpretation for a
	given assertion,

	\item the ability to formalise and reason about contexts,

	\item reasoning about events, and in particular, actions,

	\item the capacity of introspection,

	\item and finally, he points the issue of giving computer the right
	heuristics for decision making.

Coming from the perspective of natural language processing in situated context,
Roy and Reiter summarise in~\cite{Roy2005} what they see as the main challenges
to be tackled by knowledge representation systems: \emph{cross-modal
representation} systems, association of words with perceptual and action
categories (\emph{grounding}), modeling of \emph{context}, definition of the
right \emph{granularity} of models, integration of \emph{temporal modeling and
planning}, ability to \emph{match past (learned) experiences} with the current
interaction and ability to take into account the \emph{human perspective}.

Knowledge representation systems in robotics are directly affected by these
points, and we indeed integrate them in our typology, in slightly reformulated
ways.

\subsection{Article overview}
\label{sect|overview}

The paper is organized in three main sections.

Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
representation in robotics. It introduces briefly a range of different
approaches that do not focus on a specific field of robotics, and instead
provide insights on the wealth of knowledge representation techniques available
to researchers.

Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
systems deployed on \emph{service robots}: Section~\ref{sect|features}
introduces a extensive typology of requirements and features of knowledge
representation systems, that we use as foundation to compare and evaluate the
systems we examine, and section~\ref{sect|surveyed-systems} then presents each
of the eight systems, and underlines their specificities.

We finally summarize in section~\ref{sect|conclusion} the current approaches and
we attempt to identify several research directions that are not sufficiently
addressed by the current state of the art.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Knowledge Representation in Robotics: an Overview}
\label{sect|krsoverview}

\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

We propose to organize requirements from this imaginary cooking scenario where
numerous desirable features for a knowledge representation system for robotics
have been identified without any particular order.

This section proposes a more formal typology and a nomenclature of analysis
axis for a such a system. For each feature, we provide a short definition along
with links to relevant literature.

As previously mentioned, we propose six main categories, that cover both basic
and practical aspects: the formal
\taxon{A}{Expressiveness} of the system, the breadth and nature
of the \taxon{B}{Representations} allowed by the system, the
\taxon{C}{Reasoning} capabilities, how knowledge it acquired
\taxon{D}{Acquisition}, how the system integrates with the other components of
the robot sotware architecture \taxon{E}{Integration} and finally what we call
the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
base.

The next paragraph details only briefly the exact meaning of these categories
and introduce accordingly sub-categories. We kindly refer the interested reader
to~\cite{lemaignan2012symbolic} for in-depth discussion of each of these
dimensions of knowledge representation systems, as well as related literature.

Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
the main contributions of each of the surveyed KRS with respect to these
dimensions.

>>>>

\newcommand{\taxon}[2]{%
    \textbf{[#1] \emph{#2}}
}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:} #1}}

@@

The aim of this article is two-fold: We first propose a typology of
knowledge representation systems for service robotics that covers intrinsic
features, such as expressiveness or reasoning techniques, as well as extrinsic
features, such as knowledge acquisition and integration into pre-existent 
robotic architectures. We base the typology on an initial typical scenario of 
service robotics.

We then survey current knowledge representation systems in the community to
identify which knowledge-bound cognitive abilities are today well understood
and implemented, and which ones require more research efforts.

We hope this article may bring a clearer picture of the challenges of knowledge
representation and manipulation to the robotic research community.

@@

\todo{MT: For a T-RO paper, we should maybe start a bit differently and 
      emphasize the relevance for robotics and the timeliness of such an 
      overview more?}

\todo{Other ideas:}

\begin{itemize}
  \item Check if/which systems are available as open-source software?
  \item KnowRob: add more recent developments like reasoning about 3D CAD models
  \item 
\end{itemize}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Knowledge as the Root of Cognitive Abilities}
\label{sect|cognitive-abilities}

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Foundations of Artificial 
Intelligence}~\cite{Levesque2008}, Levesque reminds the reader of the manifesto 
they wrote together in 1998:

@@

    \end{itemize}
    \noindent
    and so on. With respect to robotics, our goal (like that of many in AI) is
    \emph{high-level robotic control}: develop a system that is capable of
    generating actions in the world that are appropriate as a function of some
    current set of beliefs and desires.''

To be answered, those questions have one prerequisite in common: The robot must
be able to \emph{explicitly} manipulate knowledge, and hence, needs a way to
represent the knowledge.
% 
This survey focuses on this knowledge representation issue: We first aim at
establishing a comprehensive typology of representational needs for
robotics in the specific context of service robotics, and then at painting the
current landscape of approaches in the research community in order to summarize 
the major trends and to identify possible shortcomings.

We have classified the features of knowledge representation systems (KRS) into
six broad categories: \emph{What can be represented?} or the
\emph{expressiveness} question ; \emph{How are things represented?} ; \emph{Which
reasoning techniques are offered?} ; \emph{How is knowledge acquired and
anchored?} ; \emph{How is the knowledge made available and used by other
components of the robot?} and \emph{Which actual knowledge is stored?} or the
\emph{knowledge instantiation} question.

We have also tried, for each project, to make clear \begin{inparaenum} \item
how information from perception or interaction is turned into knowledge,
\item how, in return, symbolic concept are \emph{anchored} into the robot's
sensori-motor space, \item how the knowledge representation system integrates
with the decisional layers of the robot, and how this can improve or ease
robotic control.\end{inparaenum}
\fxfatal{Do we really do it?}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\paragraph{What do we call ``knowledge''?}
\label{sect|on-knowledge}

Since we will discuss the concept of knowledge in the context of robotics in 
depth over the course of the the following pages, it is useful to make our 
terminology explicit first.
% 
Be it in philosophy, cognitive sciences or computer sciences, reaching an
agreement on a definition of ``knowledge'' seems difficult.

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, and \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language and semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the article, give a second look to
this distinction.

In our robotic context, we define knowledge as a narrower concept, while
keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
interrelated logical facts that are meaningful to the robot executive
controller}. By \emph{meaningful} we mean that they can possibly be interpreted 
to lead to a purposeful action. We will see that our main challenge while
designing a cognitive architecture is furthermore to make this knowledge as
\emph{explicit} as possible.

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question (with D,I,K 
and W standing for ``data'', ``information'', ``knowledge'' and ``wisdom'', 
respectively). In this article, we will associate ``data'' with low-level 
material like raw sensor output, and ``information'' with uncontextualised 
symbolic facts.
To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, coords)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot, would be proper knowledge.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection*{Survey Inclusion Criteria}
\label{sect|inclusion-criteria}

Every robotic system includes, explicitly or not, some kinds of knowledge 
representation systems. They may range from a simple state vector to an 
explicit symbolic knowledge base. This survey focuses on the right end of 
this spectrum: symbolic systems, suited for abstract reasoning.
% 
Besides, we have decided to restrain the set of systems to those actually
implemented on robots, and used in semantically rich environments (\ie dynamic,
partially unknown environments with a large range of different entities which
may have interactions). A typical scenario that would involve such robots is
a service robot in a human-friendly environment like a kitchen.
% 
We have further limited ourselves to systems that
\begin{inparaenum} 
  \item  run on a \emph{service robot} (that is, robots that interact with 
  objects in a semantically rich environment primarily designed for humans),
  \item  ground the knowledge in the physical world (physically embedded
  systems able to assess their environment),
  \item  are able to merge different knowledge modalities,
  \item  are able of on-line, dynamic knowledge acquisition and reasoning 
  (\ie not simple static databases).
\end{inparaenum}

\todo{MT: I think we should rethink this decision. DyKnow is a knowledge 
      processing system running on physical robots, and there's also some
      work by Lakemeyer etc in Aachen on using Golog-like things on household 
      robots}
These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.
% 
We have also chosen not to include the GOLOG language and its
derivatives~\cite{levesque1997golog, Ferrein2008, Gspandl2011} in this survey. While
several implementations on robots, including service robots, do exist, the
focus of this language is on representation and reasoning about actions and
situations, and the link with symbolic, abstract knowledge is not explicit.

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
GLAIR~\cite{Shapiro2009} or {\sc ACT-R} do have declarative knowledge
modules~\cite{Derbinsky2010} and have recently been used on service robots (see
 for instance {\sc ACT-R/E}~\cite{Kennedy2009}), they are also absent from this
survey because we did not find many references in the literature on knowledge
manipulation and representation applied to real-world robotic scenarii for
these architectures.
% 
A comprehensive reference on (bio-inspired) cognitive architectures is
available from the BICA Society~\cite{BicaCogArch2011}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Previous work}
\label{sect|evaluation-literature}

The typology we propose is partly based on a comprehensive synthesis of
classifications and analysis found in the literature. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation in the
context of service robotics.

In their chapter on Cognitive Robotics, Levesque and Lakemeyer~\cite{Levesque2008} 
present several characteristics of knowledge representation systems
for robots, stressing the need of representing the \emph{dynamics of the
world}.  Sensing is included in the knowledge representation via
\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
distinct parallel mental models; action representation and reasoning about
tasks is discussed in the context of \emph{situation calculus}; \emph{open
world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
robot programming and knowledge representation can be related. We integrate
most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define a
\emph{knowledge processing middleware} as a system supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context-dependent processing at many different levels of abstraction''. They
identify six characteristics: The system must be able to \emph{merge
information} from different, possibly distributed sources, it should support
quantitative as well as qualitative processing of information, it should offer
\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
(which requires what we here call \emph{non-monotonicity}), and finally support
\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
specification of the processing functionalities'').

Several surveys compare general cognitive architectures \cite{Langley2006,
Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
distinguish nine capabilities: recognition and categorisation, decision making,
perception and situation assessment, prediction and monitoring, planning,
reasoning, execution control, interaction and
learning/remembering/introspection. They also separately identify four
\emph{properties} of a cognitive architecture that categorize how knowledge is
handled by the architecture: representation of knowledge, organization of
knowledge, utilization of knowledge, and acquisition and refinement of
knowledge. This categorization had a notable influence on our typology, and
many of these categories are also present in our proposal.

Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, with a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology}
(how the structure of the world is captured in a representation and shared),
the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.

At an even broader scope, several authors from fields that are connected to
robotics have previously listed desirable features of artificial systems aiming
at rich cognitive abilities. For instance McCarthy recently listed the 
challenges he identified on the road to a \emph{human-level AI}~\cite{McCarthy2007}.

	\item The ability to \emph{"operate successfully in the common sense
	informatic situation"},

	\item The necessity of relying on mathematical logic, as the most fruitful
	formalism for machine intelligence,

	\item The ability to deal with \emph{approximate concepts and approximate
	theories} (that would include representing them, and reasoning with them),

	\item Non-monotonic reasoning,

	\item What McCarthy calls \emph{Elaboration Tolerance}: the ability to
	extend \emph{on demand} the closed domain of interpretation for a
	given assertion,

	\item The ability to formalize and reason about contexts,

	\item Reasoning about events, and in particular, actions,

	\item The capacity of introspection,

	\item and finally, he points at the issue of giving computer the right
	heuristics for decision making.

@@

Coming from the perspective of natural language processing in situated context,
Roy and Reiter summarize in~\cite{Roy2005} what they see as the main challenges
to be tackled by knowledge representation systems: \emph{Cross-modal
representation} systems, association of words with perceptual and action
categories (\emph{grounding}), modeling of \emph{context}, definition of the
right \emph{granularity} of models, integration of \emph{temporal modeling and
planning}, ability to \emph{match past (learned) experiences} with the current
interaction and ability to take into account the \emph{human perspective}.
% 
Knowledge representation systems in robotics are directly affected by these
points, and we indeed integrate them in our typology, in slightly reformulated
ways.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Article overview}
\label{sect|overview}

The paper is organized in three main sections.
% 
Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
representation in robotics. It introduces briefly a range of different
approaches that do not focus on a specific field of robotics, and instead
provide insights on the wealth of knowledge representation techniques available
to researchers.
% 
Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
systems deployed on \emph{service robots}: Section~\ref{sect|features}
introduces a extensive typology of requirements and features of knowledge
representation systems, that we use as foundation to compare and evaluate the
systems we examine, and section~\ref{sect|surveyed-systems} then presents each
of the eight systems, and underlines their specificities.
% 
We finally summarize in section~\ref{sect|conclusion} the current approaches and
we attempt to identify several research directions that are not sufficiently
addressed by the current state of the art.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section{Knowledge Representation in Robotics: an Overview}
\label{sect|krsoverview}

Representations for a robot's knowledge have been investigated for decades, 
starting with the seminal work on the Shakey~\cite{nilsson84shakey} robot --
which already had an internal world model based on predicate logics. Shakey
however lived in a very simple artificial world consisting of a few blocks
that could rather easily be described in logical form. Scaling towards more
complex settings required research in both AI and robotics, and in the 
following decades, these two communities operated without much interaction.
Today's robots are now able to operate in much more complex environments 
compared to Shakey's blocks world, have perception techniques that can handle 
the complexity of real scenes, manipulation capabilities that allow them to 
interact with objects, and massively more computational resources. On the one 
hand, this makes robots an interesting application for AI research again. On
the other hand, roboticists strive for more flexibility, adaptability and in 
general more semantic representation and therefore consider AI methods again. 
Besides the large-scale knowledge representation systems that will be discussed 
in the following sections, there have been many approaches to using knowledge
in parts of the robotic system.

One field that has seen much progress in the past years is the creation and 
representation of environment models, often termed ``semantic maps''. Advances 
in object recognition allowed to identify single objects in a map, which 
suggests to also store higher-level semantic information about them. The 
term ``semantic'' is however used for a range of things -- from a mere 
classification into different parts and object types~\cite{Rusu08RAS2}or the 
co-occurrence of objects in scenes~\cite{vasudevan08} up to environment 
representations in description logic~\cite{zender2008conceptual}, statistical 
relational environment models~\cite{limketkai05relational} and an embedding 
of spatial information into an encyclopedic and common-sense knowledge 
base~\cite{tenorth10envmodel}. The use of semantic maps as knowledge source 
for task planning has been investigated by~\cite{galindo08taskplanning}. 
% 
Another area of research that often includes some knowledge representation
aspects is natural-language interaction. For understanding spoken dialog
about things in its environment\cite{Mavridis2006} and for following spoken
directions \cite{kollar10natural,matuszek12commands,duvallet13imitation}, a 
robot needs to be able to ground the symbols (i.e. words) in percepts and 
actions. The natural-language text may be the result of direct interaction 
with a human, but can also be obtained from resources on the Web, for example
for mining object locality knowledge~\cite{zhou12webmining} or extracting 
task instructions from Websites~\cite{tenorth10webinstructions}.
% 
Recently, learning from experiences and ``life-long learning'' have
become popular topics in robotics. The goal is to record rich episodic 
memories during task execution~\cite{niemueller2012iros}, to reason about 
them later to extract more generic rules~\cite{winkler13memory} and to learn
generalizes task models from the collected experiences~\cite{rockel13ontology}.

\todo{add something on the IEEE standard draft for a robotics and automation ontology}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

% We propose to organize requirements from an imaginary cooking scenario where
% numerous desirable features for a knowledge representation system for robotics
% have been identified without any particular order.
% 
This section proposes a formal typology and a nomenclature of analysis 
dimensions for a knowledge representation system. For each feature, we provide 
a short definition along with links to relevant literature. As mentioned 
earlier, we propose six main categories that cover both basic and practical 
aspects: The formal \taxon{A}{Expressiveness} of the system, the breadth and 
nature of the \taxon{B}{Representations} allowed by the system, the
\taxon{C}{Reasoning} capabilities, its methods for knowledge 
\taxon{D}{Acquisition}, the \taxon{E}{Integration} of the system with other 
components of the robot software architecture, and finally what we call
the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
base.

The next section details only briefly the exact meaning of these categories
and, accordingly, introduces sub-categories. We kindly refer the interested 
reader to~\cite{lemaignan2012symbolic} for an in-depth discussion of each of 
these dimensions of knowledge representation systems as well as related 
literature.
% 
Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
the main contributions of each of the surveyed systems with respect to these
dimensions. \todo{MT: Can we list this taxonomy of capabilities as one of the 
contributions of this article?}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Representational expressiveness}
\label{sect|expressiveness}

This first axis of analysis is the intrinsic expressive power. It answers the
question of what can possibly be represented. In case it explicitly exists, the
\emph{language} of representation obviously plays an important role here.
% 
We subdivide the category \taxon{A}{Expressiveness} into five subcategories:
\taxon{A.1}{Logical formalism} classifies the underlying logical formalism of the
KRS (Horn clauses like for Prolog-based systems, Description
Logics~\cite{Baader2008}, Bayesian logics, etc.). \taxon{A.2}{Expressive Power}
quantifies the design choices made with respect to the actual expressive power
of the logical formalism: tradeoffs have to be found to allow both for high
expressiveness and tractable and decidable representations, and KRS authors
often add constraints on their representation logics to this end.
\taxon{A.3}{OWA/CWA} reflects the design choice between the \emph{open world}
(OWA) versus the \emph{closed world} (CWA) assumptions, which refers to the
possibility of representing \emph{unknown} facts. \taxon{A.4}{Uncertainty}
relates to the ability of the system to represent both the \emph{intrinsic}
uncertainty of given facts, and the \emph{extrinsic} uncertainty due to
imperfect perception by the robot. \taxon{A.4}{} relates only to the possibility
of \emph{representation}. Reasoning on uncertain facts is discussed in
\taxon{C.4}{}. The last feature of a KRS related to representation is
\taxon{A.5}{Meta-cognition} that conveys the capability for a system to
represent and exhibit it own internal structures and representation mechanisms.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Representation techniques}
\label{sect|higher-level-domain-representation}

This category focuses on questions that involve representational challenges 
(time, space, context) or require specific cognitive capabilities (theory of 
mind, introspection, memory).
% 
The dimension \taxon{B.1}{Roles}, with its sub-categories \taxon{B.1.1}{Space},
\taxon{B.1.2}{Time} and \taxon{B.1.3}{Actions}, deals with the approaches that
the KRS suggests to represent these three important \emph{roles} in robotics
(spatial relations, time and representation of actions). To give a few examples
(in no way exhaustive), \emph{spatial relations} cover techniques like
topological maps, allocentric and egocentric frame projections. \emph{Time
representation} includes Allen's interval algebra~\cite{Allen1984}, Ghallab's
chronicles~\cite{Ghallab1996} or the concept of \emph{fluent}. \emph{Actions}
are concerned with representation of \emph{pre- and post-conditions},
\emph{thematic roles} and \emph{plans}.

\taxon{B.2}{Context} describes if and how a given KRS represents the
\emph{interpretation frame} of the situation and its own knowledge. This is
often related to the initial \emph{common-sense knowledge} available to the
robot. \taxon{B.3}{Modality} refers to the ability of the KRS to represent
possible alternate worlds or belief states. A typical application of a
\emph{modal} knowledge model is the implementation of a \emph{theory of mind},
\ie an independent belief model for other agents the robot interact with.
\taxon{B.4}{Introspection} covers both \emph{self-knowledge} (what do I know
about myself?) and a weaker version focused on the explicit representation and
reasoning on the \emph{robot's own skills} (can I execute a given task?).
Finally, \taxon{B.5}{Memory} refers to the implementation of
different kinds of (bio-inspired) memory models like \emph{declarative},
\emph{procedural} or \emph{episodic}, and the attached processes
(\emph{forgetting}, \emph{selectively remembering}, etc.).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Reasoning techniques}
\label{sect|reasoning}

As one can expect, many reasoning techniques born in artificial intelligence
have found their place in robotics over the years. We propose to consider ten
subcategories of \taxon{C}{Reasoning} that reflect features we consider 
important with respect to the general requirements of robotics (like tractable 
decision making in pseudo-real time, dealing with partially unknown or uncertain 
domains, or reasoning related to interaction with the physical world). Again, 
the descriptions of these ten subcategories hereafter are very superficial. 
Readers are referred to~\cite{lemaignan2012symbolic} for details.
